Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-06 culled.tex
Updates from ShareLaTeX

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 %\usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}

\usepackage[running]{lineno}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath} 
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

\definecolor{darkgreen}{rgb}{0,0.3,0}

\usepackage[table]{xcolor}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}
\newcommand{\G}{\cellcolor{green}}
\newcommand{\Y}{\cellcolor{yellow}}

\definecolor{MyDarkBlue}{rgb}{0,0.08,0.45} 
\newenvironment{changed}{\par\color{MyDarkBlue}}{\par}

\newcommand{\ADD}[1]{\textcolor{MyDarkBlue}{{\bf #1}}}
\newcommand{\addit}[1]{\begin{changed}\input{#1}\end{changed}}

\usepackage{color}
\usepackage{listings}
\usepackage{setspace}

\definecolor{Gray}{gray}{0.9}
\newcommand{\kw}[1]{\textit{#1}}
\newcommand{\quart}[4]{\begin{picture}(75,6)
  {\color{black}\put(#3,3){\circle*{2.5}}\put(#1,3){\line(1,0){#2}}}\end{picture}}
% New Commands

\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,1}
\definecolor{Comments}{rgb}{0.5,0.5,0.5}
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{1,1,1}

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=#1,
numberstyle=\scriptsize,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
%
}}{}

\title{Learning Actionable Analytics 
 (with applications for reducing defects and reducing runtimes)}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
Rahul Krishna, Tim Menzies, Xipeng Shen\\
       \affaddr{Computer Science, NcState, USA}\\
       \{i.m.ralk, tim.menzies, xipengshen\}@gmail.com
% 3rd. author
 % use '\and' if you need 'another row' of author names
% 4th. author
\and
 Andrian Marcus\\
       \affaddr{Computer Science, UtDallas, USA}\\
       {amarcus@utdallas.edu} }

  \pagestyle{plain}
\begin{document}
  \maketitle

  % make the title area

  \begin{abstract}
 As business users demand more insightful
 analytics, data scientists need to change
 their tools. Instead of merely predicting 
 some result, they also need tools that generate ``policy advice'';
 i.e. suggestions on  how to change values in order to
 improve on some predicted value.

 This paper proposes one such policy
 generator. ``HOW'' is a near-linear-time
 tool for proposing changes to independent
 variables in order to improve on 
 the predictions of the dependent variables. Unlike other approaches
 for learning optimizations to software does not require
 an underlying of the domain (rather, HOW automatically
 learns that model as part of its operation).

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, the improvements generated by HERE can reduce
 defect counts as well as runtimes by up to
 50\% of their original values.
  \end{abstract}
  \begin{IEEEkeywords}
Instance-based reasoning, defect prediction, random forest, optimization.
  \end{IEEEkeywords}

\section{Introduction}
Business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In these two points, we say that a ``plan'' is a  set of proposed changes in order to better achieve some goal. 
Generating {\em plans} is a  very different task the standard
data mining tasks of {\em predicting} some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while planners 
  tell us {\em what to change}. For the kind of business user
discussed above, planners are the preferred to mere prediction systems
since, as one gruff user once shouted at us:
\begin{quote}
{\em ``Don't tell me what \underline{is};
tell me what to \underline{change}.''}
\end{quote}
In order to satisfy the demands of this new kind of business user, we  study the generation of {\em plans} in the context
of software analytics.  
This paper presents
HOW, a novel  {\em instance-based planner} that runs in near-linear time.
In the experiments shown below, we see that HOW's plans can
significantly reduce the expected
value of:
\bi
\item The software defects found in  twelve object-oriented JAVA systems explored in 2010 by Jureczko et al.~\cite{jureczko10};
\item The runtime of the six software systems whose configurations
are explored in 2013 by Seigmund et al.~\cite{sven12};
\ei
In summary, this paper makes six contributions that are significantly different to
prior work:
\be
\item Many software analytics papers use data to build {\em predictors} for (e.g.) defects and runtimes.
However, aside from work by the authors of this paper,
none of those papers then go further to use that data to build {\em planners}.
\item This paper offers a general framework for augmenting a prediction system such that
it becomes a planning system;
\item This paper offers a  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item This paper offers a general verification rig for assessing the value of plans generated within this rig.
\item The HOW  instance-based planner of this paper has not been presented before.
\item The experiments of this paper show that HOW planner is useful for real-world tasks;
e.g. reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\ee
The rest of this paper is structured as follows.
HOW's instance-based reasoning  makes its conclusions by
studying the space of examples. As discussed in \tion{mm},
this is a very different approach to
{\em model-based planners} that use the model
as sub-routine within a Monte Carlo simulation~\cite{me07f} or the evolutionary 
programming methods explored by the search-based software
engineering community~\cite{krall14,harman12dec}.

\section{HOW: Motivation and Methods}\label{sec:mm}

 \subsection{Motivation: Why ``HOW''?}
HOW was inspired by certain  limits observed in 
the GALE optimizer~\cite{krall14}. 
A well-known effect in machine learning (and, in computer vision
and compression~\cite{levina04}) is that the $n$ dimensions of a data set can be represented by a  much
lower-dimensional approximation. GALE finds those $m << n$ dimensions then mutates examples
towards the better end of each dimension $m_i$. Since it explores $m \ll n$ dimensions,
GALE terminates much faster than standard optimiser (e.g. for one large model, 4 minutes versus 7 hours~\cite{krall14}).
Better yet, GALE's rapdily generated  solutions are competitive (or better) than those found by 
slower optimizers.

A limitation with GALE is that it makes extensive use of
an underlying domain model to asses generations of newly mutated instances.  This is an issue since
in some domains, no such model exists or is readily accessible. For example, this   paper
began when, after briefing a business client on GALE, they asked:
\begin{quote}
{\em ``But what if I do not have a model of my business?''}
\end{quote}
This is a legitimate concern.
Building and certifying a model  can  takes  long time, especially in software engineering. For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}).  This is not an isolated example. One of us (Menzies) teaches a graduate class in automated model-based software engineering. 
Students in that class are asked to find (or build) a non-trivial model of
some software process (where ``non-trivial'' means something that processes
a  model at least an order of magnitude more complex than Brooks' Law). 
While the  number of such models has been growing in recent years\footnote{See the
catalogue at goo.gl/nv2AVK.}, 
the total number remains very small-- which is an observation consistent with
SE models typically require  much effort to build and certify.

Also, even when there is an existing model, instance-based reasoning can be useful.
Models need maintenance lest they become out-dated. Elsewhere, we have described our
extensions to the USC models to enable reasoning about agile software developments. 
It took many months to implement and certify those extensions~\cite{me09i,me09j}.
The problem of model maintenance is another
motivation to look for alternate methods that can be automatically updated whenever new data arrives.

Finally, sometimes  it is easier to access instances of a model's behaviour than the model
itself. For example, in prior work with Martin  Feather from the Jet Propulsion 
Laboratory~\cite{fea02a},  our research partner could not share a
propriety model from within the NASA firewalls. However, they could share 
logs of the input/output of that model.

Note that this lack of access to a workable model is not just fatal for GALE, but also for every
other model-based optimization algorithm such as simulated
annealling or evolutuionary methods such as   
NSGA-II, NSGA-III, SPEA2, IBEA, PSO, DE, MOEA/D, etc.~\cite{deb00a,zit02,zit04,%
deb14,Cui2005a,storn97,zhang07:TEC}. Accordingly, this paper seeks
an optimization methods that avoids the problem of models that are missing, slow
to build, hard to maintain, or difficult to access.

\subsection{Methods: How to Build ``HOW''?}

A general model for {\em predictors} is as follows. 
During testing, each test instance gains a prediction 
by finding which region of the data it falls into.
During training, those regions are inferred from
divisions that ``chunk'' together regions with instances
with similar properties. These dividers can be categorical (as in the case of decision trees) or probabilistic (as in the
case of Naive Bayes) or learned via some distance minimization
process 
(as in the case of support vector machines) or generated
directly by some clustering algorithm.

Those same regions can be used for instance-based 
{\em planning}-- but instead of leaving test instances in the regions
found during training, we displace them towards regions
with better objective scores. 
Two important design decision with HOW 
is how to determine the {\em magnitude} and {\em direction} of the
displacement.

For HOW v1.0 (reported in this paper), we committed to a policy
of {\em minimal displacement}  since we do not wish to alarm developers and managers
with large scale change requests.
HOW's minimal displacement policy is to generate
many small clusters, then limit displacement to the distance
to the nearest cluster.
Note that minimal displacement makes HOW very different
to evolutionary optimizers.
Those algorithms, which include  
GALE, NSGA-II, NSGA-III, SPEA2, IBEA, PSO, DE and  MOEA/D~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,storn97,zhang07:TEC} can make large scale
displacements to the original population.

For the purposes of clustering, we use the WHERE recursive
descent procedure of \fig{where} that converts
$N$ instances into $\sqrt{N}$ clusters.
WHERE is a spectral clusterer~\cite{kamvar03}; i.e. it
executes in near linear time to generate
cluster using dimensions  discovered
by a  PCA-like analysis (PCA=principle components analysis).

\begin{figure}[t!]
\small
\begin{tabular}{|p{.95\linewidth}|}\hline
WHERE recursive divides data as follows.
It finds the two   distance instances,  $X,Y$
by picking any instance $W$ at random, then setting $X$ to its most
distant instance, then setting $Y$ to the instance most distant from
$X$. 
Note that finding $X,Y$ requires only $O(2N)$ comparisons
of $N$ instances.

Let $c=dist(X,Y)$. Using $c$, we next  project each instance $Z$
onto a line that  runs between $X,Y$ as follows. If $a=dist(Z,X)$ and
$b=dist(Z,Y)$, the $Z$'s
distance from   {\em X} along the $XY$ line is
$x=(a^2 + c^2 - b^2)/(2c)$.
To recurse, split the data at the median $x$ value of all instances then call WHERE on each half (stopping when
one half has less  than $\sqrt{N}$ of the original population

In the above, to computer distances, we use
the Euclidean measure recommended for
instance-based reasoning by Aha et al.~\cite{aha91};
i.e. $\sqrt{\sum_i(X_i-Y_i)^2}$ where $X_i,Y_i$
are values normalized 0..1 for the range min..max.
\\\hline
\end{tabular}
\caption{Clustering data with WHERE. From~\cite{me12d}.}\label{fig:where}
\end{figure}

Having  limiting the displacement {\em distance} to no further that
the nearest  cluster,
there is still the issue of how to select the displacement 
{\em direction}. 
To determine that direction, we reflect on the objective
scores of each cluster.
For example:
\bi 
\item
For the Seigmund data,
the objective score is the runtime of the test suite of some compiled program;
\item
For the Jureczko data, the objective score
is a boolean that is true if defects have been noted
for that class.
\ei
We say that each cluster can be represented by one
{\em exemplar} that
holds independent values that best characterize that cluster:
\bi
\item
For
instances with numeric objectives (e.g. the Seigmund data), 
the exemplar of a cluster is the instance
with the best objective score.
\item
For instances with discrete objectives (e.g. Jureczko data),
there is less differentiation between the objective scores.
Hence, for this kind of the data, the exemplar of a cluster is
an instance synthesized from the median values of the instances
in that cluster.
\ei
Given the above, we can now define a {\em plan} to be
a set of values to be added to a test instance in order
to reduce the distance between that instance and the exemplar: 
\[Y_j = Y_j + f*(X_j - Y_j)\]
In the above, $Y_j$ is an independent variable within the
test instance $Y$; $X$ is the exemplar found at training time;
and  $f$ is a constant that controls
how far we  move an instance (which is tuned for different data
sets).

A core design principle for HOW was ``minimal change'';
i.e. make very small changes to the current instance.
\ei
There are many ways to find those regions 
When designing HOW, our premises were:

This is no domain model; (b)~we have some
labelled data and (b)~we seek minimal
disrupttion
HOW divide training into regions where the training
data has similar properties. Each region is then examined and divided into ``west'' and ``east''
where the region's ``west'' end has examples with worst performance stores than the east
end (mnemoc: west starts with ``w'' and is short for ``worst' and east starts with ``e;;]which is shirt
for excellent).so the ``e'' in

\be
\item Take  a population $P$ of examples
\item Separate all $p \in P$ examples into their $n$ independent variables and 
their $o$ scores;
\item Find $m \ll n$ lower dimensional approximations for the $n$ independent variables;
\item Look for the worst and best ends of each $m$  dimension (where the better ends
have better objective scores);
\item Find the $q \subseteq P$ subset of the population with ``better'' objective scores;
\item Generate the next population $P'$ by adjust the independent variables
of the $q$  examples
by a random amount (in the direction of  the  better ends of each dimension).
\item Apply some model to discover all the objective scores of $P'$
\item Set $P=P'$ and go to step 1
\ee

PCA explores the $e$
eigenvectors of the correlation matrix between $d$
variables~\cite{pearson1901}.  
This is a useful approach since, usually,
only a few $e \ll d$ eigenvectors are required
since multiple inter-related variables $i,j,k \subseteq d$ can be represented
by a single  eigenvector.
Also, noisy variables from $d$ are
ignored since they  do not contribute to the signal in the data.
Further,
redundant  variables  become (approximately) parallel lines
in $e$ space. For all such redundancies \mbox{$i,j \in d$}, we
can ignore $j$
since effects that change over $j$ also
change in the same way over $i$.

Algorithms like PDDP~\cite{boley98} use PCA to
recursively cluster data into smaller regions. Each
step of that process requires an $O(N^2)$ analysis
of the data to compute the components~\cite{ilin10}. Worse, it requires an evaluation of every example.
Using WHERE, competitive results can be achieved
using an $O(2N)$ analysis.
WHERE uses a heuristic proposed by
Faloutsos and Lin~\cite{Faloutsos1995}
(which
Platt has
  shown computes a Nystr\"om
  approximation to the first component of
  PCA~\cite{platt05}).
Initially, WHERE inputs
$N$
instances $N_1,N_2,..$. Secondly, WHERE
picks any
instance $N_i$ at random. Thirdly, WHERE
finds the instance  {\em West}~$\in N$ that is
furthest from $N_i$. Finally,
if funds the point {\em East}~$\in N$ 
that is furthest from {\em West}. To computer these distances,
we use the standard Euclidean measure recommended for
instance-based reasoning by Aha et al.~\cite{aha91};
i.e. $\sqrt{\sum_i(x_i-y_i)^2}$ where $x_i,y_i$
are values normalized 0..1 for the range min..max.

\be
\item Clustering the training data 
\item 
(c) finding which end of the cluster has better performance scores.
(d) find
\ee

\section{Methodological Comments:\\ from ``Prediction'' to ``Planning''}
A major  problem with building a planning mechanism in software engineering is ``how to verify the generated
plans?''. For example,  in the experiments shown below,  HOW comments  on how to reduce
software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be mnaually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}. How are we to verify that HOW's recommendations actually work?

Some organizations have the resources to 
run repeated trials to assess the effect of different treatments.
For example, in one remarkable recent study, Bente et al. report results
were the same specification was developed into a product by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to $N$ different teams, apply
HOW's plans the code used by  $N/2$ teams, then check in 12 months time
if those halves of the teams have fewer defects than the other.

Very few industrial or research groups have access
to the kinds of resources using by Bente et al. (evidence: in the size years since the
publication of that work, we know of only one of two similar studies). Hence, it
we wish to verify HOW's plans, then some other approach is required that is less
resource intensive.

\subsection{Proposed Framework}

We propose using prediction systems (built by data mining) as a verification
tool for planners. In domains with data, from which data miners can build
predictors of adequate performance, then those predictors can assess the value
of the changes proposed by HOW.
As shown in \fig{work}, 
project data is divided into two disjoint sets {\em train} and {\em test}
(so \mbox{{\em train} $\cup$ {\em test} $=\;\emptyset$}).
Next, from the train set, we build both a {\em planner} as well
as a {\em predictor}. Our general framework does not   commit to any particular  choice
of {\em planner} or {\em predictor}. For the purposes of this discussion 
our {\em planner}
will be HOW (discussed below)and our
our {\em predictors} will be the Random Forest and Random Forest
Regressor\footnote{Decision tree learners find the attribute whose value's
select for for sections of the data with least variation in the class
variable. Decision tree learners then recurse on each selection~\cite{breiman84}. A random forest is a set of decision trees, each built
from some random subset of the rows and columns. The conclusions
of the forest are a ``vote'' generated by all the trees~\cite{Breiman2001}. 
A standard Random Forest predicts for a discrete class variable while
a Random Forest Regressor generates numeric predictions.} taken from the SciKit
Learn toolkit~\cite{Pedregosa2012}.

Turning now to the {\em test} data, this is passed to the {\em predictor}
to generate a baseline {\em before} performance score.
Note that if our {\em predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the {\em planner} to alter the {\em test} data;
then (2)~apply the {\em predictor} to the altered data {\em test'};
then (3)~return the performance delta.

\begin{figure}[!t]
\small 
\[
\begin{array}{r}
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{predictor}(\mathit{test})\\
           \mathrm{\bf if\;}\mathit{before} & <  \mathit{satisfactory}\\
           \mathrm{\bf then}  & \mathrm{{\bf return}\; 0}\\~\\
           \mathrm{\bf else} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} \mathit{after} /  \mathit{before}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

\caption{How to use {\em predictors} for {\em planning}.}\label{fig:work}
\end{figure}

\subsection{Are These Plans ``Causal''?}
With one additional stipulation,   this framework can be used to check if the learned
plans can be said to {\em cause} a change in the domain being explored.

In the literature, there are two main definitions of causality: Hill's H-causality~\cite{Hill1965}
and Granger's G-casualty~\cite{granger80}. 
Of the two, Hill's definition is the most strict. An effect
is H-causal if it satisfies nine criteria including
{\em  plausibility};
{\em strength} of the effect;
{\em 
specificity}, i.e. locality of the effect
to specialized groups;
{\em
support} by experimental evidence;
{\em 
coherence} between cause-effects noted
in the field and in laboratory samples;
{\em
temporally} i.e. the effect has to follow the cause;
{\em 
analogy}, i.e.  similar causes have similar effects;
{\em
biological gradient}, i.e.  more exposure to the cause leads to more of an
effect; and
{\em
consistency}, i.e.  the same cause-effect pairings have been seen by different persons
in different places.

It is hard to demonstrate H-causality in software engineering.
For example, Hill's criteria emphasises repeatability and laboratory studies-- both of which are
hard to perform in software engineering (e.g. see the continued debate about
the value of using undergraduate students to surrogates for industrial
practitioners in software engineering research~\cite{Carver2003}). Also, Hill's requirement
for consistency is far easier to achieve in studies of small closed systems (e.g. bacteria in a petri dish)
than in large socio-technical systems (e.g. open source projects). In the latter,
effects are more {\em probabilistic} than {\em categorical}-- by which we mean that 
some cause tends to cause some effect at some probability $p$ less that one.
This in turn means that at some probability $q$ (where $0 \le q < p \le 1$)
that we will find examples of {\em inconsistency} where some cause does {\em not} lead
to some effect (thus violating one of Hill's criteria for demonstrating causality).

The  difficulties associated with Hill's criteria lead Granger~\cite{granger80}
to propose another more practical criteria\footnote{Granger  himself comments that
his proposal was based on earlier work by Norbert Weiner~\cite{Seth2007}.}.
\mbox{G-causality} is based on building predictions based on {\em past} data,
then showing that this works for {\em future} data~\cite{granger80}. 
In should be noted  that the standard 10-way cross-validation
procedure used  in data mining does not necessarily demonstrate G-causality.  In a N-way cross-val experiments, the data is divided into $N$ bins.  
$N-1$ bins are used for learning and the results are tested
on the  remaining bin.
This is repeated $N$ times, each time using a different bin for the test set.
This approach can not be used to demonstrate causality since, if the initial data is collected
over (say) $N$ months, then it is possible that the test set contains observations collected
{\em before}   items seen in the training set.

@@

We adapt the work of Jiang,Lumpe et al.  as follows. In  \fig{work}, we demand that the {\em test}
data is not necessarily collected before the {\em train} data:
\bi
\item[\#1:] For data collected from software released in multiple versions, we should hence use  versions  $i,j$ to build models that are applied to version  $k$ where $i<j<k$.
\item[\#2:] For data collected from Monte Carlo simulations (where inputs are selected at random)
then as long as simulation $i$ is not used to adjust simulation $j> i$, then any item in that
data could be generated before any other. For this kind of data, we  can use the first $X\%$
of the simulation for training, and the remaining data for testing.
\ee
In the two case studies of this paper, the Jureczko object-oriented defect data comes from software
released in multiple versions while the Seigmund configuration data comes from Monte Carlo simulations.
Hence, we use procedure \#1 for the defect data and procedure \#2 for the configuration data.

With these changes,
we can reasonably claim that the plan's generating by HOW are G-causal.

\subsection{When Not to Plan: the Problem of Poor Predictors}

The clear advantage of the approach of \fig{work} is that it can be readily and automatically
applied to domains with historical data. That said, the obvious drawback with this approach
is that our assessment of the plans is only as good as the predictor. To 
mitigate for this {\em problem of poor predictors}, it is hence wise to demand high standards for the {\em
satisfactory} threshold used in this \fig{work}.

How prevalent is the problem of poor planners? 
In the general case, this is hard to answer. However, as discussed in this section, for the 17 data
sets explored in this study, only 
%XXX 
3 suffer from this problem.

This study's data from two types of domains: the Jureczko   object-oriented data  
and the Seigmund configuration data. 
For the Seigmund data, the plans generated by HOW advise on how  to change the   configurations settings within a Makefile
(in order to make resulting compiled
code  run faster). Here, performance can be measured in terms of  difference
between the predicted runtime $p$ of test case items and their actual runtimes $a$
using  $s= 100*(1- \frac{abs(a - p)}{a})$.
Using engineering judgement, we define satisfactory $s$ as anything over   70\%. 
This paper  explores    six Seigmund configuration data sets:  Berkeley DB (Java and Windows versions), Apache, SQLite, LLVM, and
  x264\footnote{Available from the performance prediction section of PROMISE
  http://openscience.us/repo/performance-predict.}. 
  After splitting that data into equal sized train:test groups, a Random Forest
  Regressor trained on one half, then applied to the other, achieved nearly perfect scores of
%issue 24
\[s=\{99.9, 99.8, 99.4, 99.1, 96,1\}\]
(Aside: if the reader doubts that such high scores are achievable, we note that these scores are 
consistent with those achieved by predictors built by Seigmund et al.~\cite{sven12}.)

\begin{figure}
\input{deltas1}
\caption{Jureczko data sets, experiments with tuning.}\label{fig:jtune}
\end{figure}
For the  Jureczko   object-oriented data,
the plans generated by HOW advise   how to change static code attributes in order to reduce defects in
Java classes.  Here,
performance can be measured in terms of recall (percent of faulty classes in
the test data detected
by the {\em predictor}). Using engineering judgement, we define satisfactory recall as anything over   70\%. 
This paper explores 12 of the Jureczko   object-oriented static code data sets: Ant, Camel, Ivy, Jedit, JBeans, Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository openscience.us/repo/defect/ck.}.
For each of these data sets, we generated  {\em before} scores  by finding the latest
version $k$, then training  a predictor on versions $k-2$ and $k-1$. As reported below,
our initial recall scores from this data were unpromising (very low recall values). However,
after applying certain data pre-processing steps (described below), a Random Forest generated
on the $k-2,k-1$ versions achieved recalls on version $k$ of
%issue24
\[\mathrm{recall}=\{98,88,86,84,82,81,79,78,78,\overbrace{55,23,20}^{\mathrm{JBeans,Poi,Xalan}}\}\]
From the above, we conclude that all the Seigmund configuration data and most of the Jureczko data
avoids the problem of poor predictors. Hence, we can use that data for the planning
experiments described below.

\section{Experimental Methods}

blah blah data. algorithms. statistics

\section{Releated Work}

Our prior work in generating policies relied upon a Monte Carlo
analysis of an underlying model~\cite{me07f}. This approach had three major problems.
Firstly, enough domain knowledge had to be collected to build that model
and this can take a long time. For example, in our previous work we used three software process models from
the Unviersity of Southern California that were initially
developed and updated for nearly two decades from 1981~\cite{boehm81} to 2000~\cite{boehm00b}.

Secondly, there is the computational cost of running those models.
While

~\cite{}-- which suffered from the problem of (a)~the effort required to collect the domain knowledge to build
problem in certifying the underlying model;
 (b)~the computational problem of running that model enough
times to learn patterns in that data; and (c)~ requrearea used large scale simulations or large

While
our case studies are defect and runtime reduction, HERE is a general
purpose algorithm that could be applied to any table of data with at least
one column whose values can be sorted from ``desired'' to ``undesired''.

{\bf RQ1: Can Instance-based Learning techniques be used to create Planning tools that can better analyse software defect?}

{\bf RQ2: Can lessons learnt from local policy generators be useful for generating solutions that can be used to mitigate defects in software systems?}

{\bf RQ3: Can such instance-based planners be applied to other software engineering paradigms?}
We check if the proposed instance-based planner is applicable to performance prediction.

\section{Motivating Example}
\section{Background}
\section{Instance Based Learning}

Assessing the quality of solutions to a real-world problem can be extremely hard~\cite{menzies2005xomo}. In several software engineering applications, researchers have models that can emulate the problem, for instance there is the COCOMO effort model~\cite[p29-57]{boehm2009software}, the COQUALMO defect model~\cite[p254-268]{boehm2009software}, Madachy’s schedule risk model~\cite[p284-291]{boehm2009software}, to name a few. Using these models it is possible to examine several scenarios in a short period of time, and this can be done in a reproducible manner. However, models aren't always the solution, as we shall see.

There exist several problems where models are hard to obtain, or the input and output are related by complex connections that simply cannot be modeled in a reliable manner, or generation of reliable models take prohibitively long ~\cite{Ludewig2003}. Software defect prediction is an excellent example of such a case. Models that incorporate all the intricate issues that may lead defects in a product is extremely hard to come by. Moreover, it has been shown that models for different regions within the same data can have very different properties \cite{localvsglobal}. This makes it extremely hard for one to design planning systems that are capable of mitigating these defects.

An alternative approach would be to make use of an instance based approach, an example of case based reasoning strategy, in place of the conventional model based approach. Instance based approaches have been used extensively by the effort estimation community, see~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}.

This approach has been proposed as an alternative to closed form mathematical models or other modeling methods such as regression \cite{keung2008analogy}. There are several other reasons for instance based approaches being a useful tool, see~\cite{6600685}. As pointed out by~\cite{walkerden1999empirical} it can be used with partial knowledge of the target project at an early stage which could be a very useful tool in preventing software defects. Instance based approaches are also rather robust in handling cases with sparse samples \cite{1438374}. All these features are desirable and suggest that instance-based approach is a useful adjunct to traditional model based approach.

In this paper the instance based learning principle is used to first cluster the training data and then create contours in the data. These contours can then be used by test cases to infer meaningful information in order to improve it. The entire process is described in greater detail in the following sections. We would like to stress that the purpose of our work is not to promote instance-based planning over model based analytics, rather it is to suggest an alternative when model based approaches cannot be used.

\subsection{Spectral Learning using WHERE}
The algorithm uses WHERE to recursively cluster the input data to identify subsets in the training data that a test instance can learn from. WHERE is a spectral learner which uses the FastMap heuristic to estimate the first principal component. It then recursively partitions the data into two halves along the Median point of the projection on the first principal component, terminating when a half has less than $\sqrt{N}$ items.

\subsection{Planning changes using HERE}
We propose the HERE algorithm as a tool to plan changes in the original test data. Figure~\ref{fig:what} highlights the procedure by which HERE operates. It begins by creating clusters which are generated using the WHERE algorithm discussed above. Following this, a nearest neighbour scheme is applied to identify pairs of nearby clusters. A contour is constructed with these clusters at the vertices to characterize the cluster pairs.

The mutation policy works by projecting the test instance onto the contours and identifying the the contour on which the test instance has the largest scalar projection. The planner then reflects over the vertices of the chosen contour, identifying the \textit{better} vertex among the two. Now, a new instance is generated by \textit{mutating the attributes} of the test instance towards the better vertex and away from the worse vertex. This process is repeated for all the test instances that are considered defective by the defect prediction scheme, discussed below.

\subsection{Feature Weighting} \label{fwt}
The algorithm also allows for a feature weighting and pruning scheme. We have employed a simple (and fast) attribute ranking method where the features are sorted based on their ability to find meaningful splits in the dependent class of the data, see ~\cite{hall03}. Features that can find splits with the \textit{lesser variance} in the dependent variable are \textit{ranked higher} than features that don't.

We reason that the most informative feature must undergo a larger change as compared to other features. Therefore, when feature weighting is employed, the higher ranked features are mutated by a larger extent. Additionally, we have also included information pruning. Here we mutate only the top X\% of the features in each test instance. The remaining features remain unaltered.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{_figs/WHAT-Clusters2.pdf}
\caption{HERE}
\label{fig:whatflow}
\end{figure}

>>>>

In should be noted  that the standard 10-way cross-validation
procedure used  in data mining does not necessarily demonstrate G-causality.  In a N-way cross-val experiments, the data is divided into $N$ bins.  
$N-1$ bins are used for learning and the results are tested
on the  remaining bin.
This is repeated $N$ times, each time using a different bin for the test set.
This approach can not be used to demonstrate causality since, if the initial data is collected
over (say) $N$ months, then it is possible that the test set contains observations collected
{\em before}   items seen in the training set.

@@

We adapt the work of Jiang,Lumpe et al.  as follows.
